Developmental trajectories of word recognition: a priming study in bilinguals

### 2023-11-27 manuscript/manuscript.qmd
Fix figure

```{r setup}
#| label: setup
#| echo: false
#| message: false
#| warning: false
# load objects
targets::tar_config_set(
	store = here::here("_targets"),
	script = here::here("_targets.R")
)

@@

targets::tar_load(
	c(
		participants,
		stimuli,
		vocabulary,
		attrition_trials,
		attrition_participants,
		bvq_data,
		gaze,
		data_bcn,
		data_oxf
	)
)

targets::tar_load(
	c(
		model_fits_bcn,
		model_loos_bcn,
		model_fits_oxf,
		model_loos_oxf
	)
)

# prepare data
attrition <- lst(
	participants,
	attrition_participants,
	attrition_trials
) |>
	reduce(inner_join, by = join_by(session_id)) |>
	unnest_wider(matches("all"), names_sep = "_") |>
	rename_with(\(x) gsub("is_valid_gaze_all_|is_valid_vocab_all", "", x))

library(knitr)
library(kableExtra)
library(ggplot2)
library(gt)
library(gtExtras)
library(patchwork)
library(english)
library(gtExtras)

# set ggplot theme and colour palette
my_theme <- theme_minimal() +
	theme(
		panel.grid = element_blank(),
		axis.line = element_line(colour = "black"),
		text = element_text(size = 12, colour = "black"),
		axis.text = element_text(colour = "black")
	)

@@

clrs <- c("#003f5c", "#58508d", "#bc5090", "#ff6361", "#ffa600")

options(
	ggplot2.ordinal.fill = clrs[c(1, 4, 5)],
	ggplot2.ordinal.colour = clrs[c(1, 4, 5)],
	ggplot2.discrete.fill = clrs[c(1, 4, 5)],
	ggplot2.discrete.colour = clrs[c(1, 4, 5)],
	ggplot2.continuous.fill = ggplot2::scale_color_gradient,
	ggplot2.continuous.colour = ggplot2::scale_color_gradient
)

@@

	cols_merge(c(age_min, age_max), pattern = "{1}-{2}") |>
	cols_merge(c(age_mean, age_sd),
			   pattern = "{1} ({2})") |>
	cols_merge(c(doe_catalan_mean, doe_catalan_sd),
			   pattern = "{1} ({2})") |>
	cols_merge(c(doe_english_mean, doe_english_sd),
			   pattern = "{1} ({2})") |>
	cols_merge(c(doe_spanish_mean, doe_spanish_sd),
			   pattern = "{1} ({2})") |>
	tab_spanner("Sample size", starts_with("n")) |>
	tab_spanner("Age (months)", matches("age")) |>
	tab_spanner("Catalan", matches("catalan")) |>
	tab_spanner("English", matches("english")) |>
	tab_spanner("Spanish", matches("spanish")) |>
	tab_spanner("Degree of Exposure (%)", matches("doe")) |>
	cols_label(n = "Sessions",
			   n_participants = "Participants",
			   age_mean = "M (SD)",
			   age_min = "Range",
			   doe_catalan_mean = "M (SD)",
			   doe_english_mean = "M (SD)",
			   doe_spanish_mean = "M (SD)") |>
	tab_style(cell_text(style = "italic"),
			  cells_column_labels()) |>
	tab_style(cell_text(style = "italic"),
			  cells_column_spanners())
```

@@

```{r fig-vocabulary}
#| label: fig-vocabulary-oxf
#| fig-cap: "Participant receptive vocabulary sizes across ages and language profiles. Vocabulary size scores from the recurrent participants are shown linked. For visualisation purposes, testing sessions were aggregated at three age points: 21, 25, and 30 months. Points and intervals indicate mean and standard deviations, respectively."
#| fig-width: 6
#| fig-height: 4
vocab_plot_data <- vocabulary |>
	inner_join(filter(attrition_participants, is_valid_participant),
			   by = join_by(session_id)) |>
	inner_join(select(participants, location, lp, 
					  child_id, session_id, age_group),
			   by = join_by(child_id, session_id)) |>
	pivot_longer(c(l1_prop, total_prop),
				 names_to = "type",
				 values_to = "prop") |> 
	mutate(location = factor(location, levels = c("Oxford", "Barcelona")),
		   lp = gsub(" \\(English\\)", "", lp),
		   side = case_when(location=="Oxford" ~ "both",
		   				 lp=="Monolingual" ~ "left",
		   				 lp=="Bilingual" ~ "right"),
		   type = if_else(type=="l1_prop", 
		   			   "L1 vocabulary", "Total vocabulary"),
		   study = if_else(location=="Oxford", 
		   				"Study 1 (English)", "Study 2 (Catalan/Spanish)"))

vocab_plot_means <- vocab_plot_data |> 
	summarise(prop_sd = sd(prop, na.rm = TRUE),
			  prop = mean(prop, na.rm = TRUE),
			  n = n(),
			  .by = c(age_group, type, side, lp, type, location)) |> 
	mutate(.lower = prop - (prop_sd/sqrt(n)),
		   .upper = prop + (prop_sd/sqrt(n)))

vocab_plot_data |> 
	ggplot(aes(age_group, prop, 
			   colour = lp,
			   fill = lp,
			   side = side)) +
	facet_grid(type~study) +
	geom_line(aes(group = child_id),
			  colour = "grey",
			  alpha = 1 / 4, linewidth = 3 / 4) +
	geom_dots(aes(side =),
			  dotsize = 1.5,
			  scale = 0.25,
			  binwidth = 0.025,
			  alpha = 2 / 4) +
	geom_errorbar(data = vocab_plot_means,
				  aes(ymin = .lower,
				  	ymax = .upper),
				  colour = "black",
				  show.legend = FALSE,
				  width = 0.15,
				  linewidth = 1/2,
				  position = position_dodge(width = 1)) +
	geom_point(data = vocab_plot_means, 
			   colour = "black",
			   show.legend = FALSE,
			   size = 1.25,
			   position = position_dodge(width = 1)) +
	# geom_label(data = vocab_plot_means,
	# 		   aes(label = sprintf("%.1f%%\n[%.1f, %.1f]",
	# 		   					prop*100, .lower*100, .upper*100),
	# 		   	y = 0),
	# 		   size = 2.5,
	# 		   fill = "white",
	# 		   vjust = 0,
	# 		   show.legend = FALSE,
	# 		   label.r = unit(0, "lines"),
	# 		   label.size = 0,
	# 		   label.padding = unit(0.1, "lines"),
# 		   position = position_dodge(1),
# 		   alpha = 1/4) +
# geom_point(alpha = 1) +
labs(x = "Age (months)",
	 y = "Vocabulary size",
	 colour = "Group",
	 fill = "Group") +
	scale_y_continuous(labels = scales::percent,
					   limits = c(0, 1)) +
	theme(panel.grid.major.y = element_line(linetype = "dotted",
											colour = "grey"),
		  legend.position = "top",
		  axis.title.x = element_blank(),
		  legend.title = element_blank())
```

@@

We created fours lists of trials, across which the same target-distractor pair appeared with a different prime, counterbalancing the condition to which it belonged. For instance, in list one the *ball*-*trousers* pair was preceded by *bike* (*Related/Cognate*), by *butterfly* (*Related/Non-cognate*) in list two, and by *star* and *nose* (*Unrelated*) in lists three and four. @tbl-stimuli shows a detailed summary of the stimuli properties, broken down by trial type and testing language. Lexical frequencies were extracted from the English corpora from the CHILDES database [@macwhinney2000childes; @sanchez2019childes] as counts per million words, and transformed into Zipf scores for easier cross-language comparison [@zipf1945meaning; @van2014subtlex].

```{r tbl-stimuli}
#| label: tbl-stimuli
#| tbl-cap: "Summary of stimuli properties by trial type."
stimuli |>
	unnest_wider(c(freq, nphon)) |>
	summarise(across(c(matches("n_phon|freq_|lv_|nphon"), duration),
					 list(	mean = \(x) mean(x, na.rm = TRUE),
					 	  sd = \(x) sd(x, na.rm = TRUE))),
			  .by = c(trial_type, test_language)) |>
	select(-matches("distractor")) |>
	mutate(test_language = if_else(test_language=="English",
								   paste0("Study 1: ", test_language),
								   paste0("Study 2: ", test_language))) |>
	arrange(test_language) |> 
	gt(rowname_col = "trial_type",
	   groupname_col = "test_language") |>
	# summary_rows(groups = c("Unrelated", "Non-cognate", "Cognate"),
	# 			 columns = is.numeric,
	# 			 fns = list(Total ~ mean(.)),
	# 			 fmt = ~ fmt_number(., decimals = 2)) |>
	grand_summary_rows(	columns = is.numeric,
						fns = list(Total ~ mean(.)),
						fmt = ~ fmt_number(., decimals = 2)) |>
	tab_spanner("Phonemes", matches("nphon")) |>
	tab_spanner("Frequency (Zipf)", matches("freq")) |>
	tab_spanner("Phon. similarity [0-1]", matches("lv")) |>
	fmt_number(is.numeric, decimals = 1) |>
	cols_merge(c(freq_prime_mean, freq_prime_sd),
			   pattern = "{1} ({2})") |>
	cols_merge(c(freq_target_mean, freq_target_sd),
			   pattern = "{1} ({2})") |>
	cols_merge(c(nphon_prime_mean, nphon_prime_sd),
			   pattern = "{1} ({2})") |>
	cols_merge(c(nphon_target_mean, nphon_target_sd),
			   pattern = "{1} ({2})") |>
	cols_merge(c(lv_pp_mean, lv_pp_sd),
			   pattern = "{1} ({2})") |>
	cols_merge(c(lv_pt_mean, lv_pt_sd),
			   pattern = "{1} ({2})") |>
	cols_merge(c(duration_mean, duration_sd),
			   pattern = "{1} ({2})") |>
	cols_label(trial_type = "",
			   freq_prime_mean = "Prime",
			   freq_target_mean = "Target",
			   lv_pp_mean = "Prime-Prime",
			   lv_pt_mean = "Prime-Target",
			   duration_mean = "Duration (s)",
			   nphon_prime_mean = "Prime",
			   nphon_target_mean = "Target") |>
	tab_style(cell_text(style = "italic"),
			  list(cells_column_labels(),
			  	 cells_column_spanners(),
			  	 cells_grand_summary()))
```

@@

After trials that matched any of the aforementioned exclusion criteria from the dataset, we excluded participants who did not provide at least two valid trials in the *Cognate* condition, the *Non-cognate* condition, or the *Unrelated* condition (*n* = `r n_exc_participants`), and participants with a L1 vocabulary size lower than 10% (*n* = `r n_exc_participants_vocab`). The final dataset included `r format(n_trials_valid, big.mark = ",")` trials from `r n_sessions_valid` testing sessions, generated by `r format(n_participants_valid, big.mark = ",")` distinct participants. Of those participants, `r n_longitudinal[1, 2]` provided data from one experimental session, `r n_longitudinal[2, 2]` provided data from two experimental sessions, and `r n_longitudinal[3, 2]` provided data from three experimental sessions. From the trials included in the final dataset dataset, `r n_trials_condition$unrelated_2$n` were *Unrelated* trials (`r n_trials_condition$unrelated$n` previously excluded), were `r n_trials_condition$non_cognate_2$n` *Related/Non-cognate* trials(`r n_trials_condition$non_cognate$n` previously excluded), and `r n_trials_condition$cognate_2$n` were *Related/Cognate* trials (`r n_trials_condition$cognate$n` previously excluded).

```{r tbl-attrition}
#| label: tbl-attrition
#| tbl-cap: "Participant-level and trial-level sample size after applying inclusion criteria. The number of excluded participants and trials is indicated between parentheses."
expanded <- expand(attrition, lp, trial_type, is_valid_trial)

n_lp <- attrition |>
distinct(is_valid_participant, lp, session_id) |>
count(is_valid_participant, lp) |>
pivot_wider(names_from = is_valid_participant,
values_from = n,
names_repair = janitor::make_clean_names) |>
mutate(n_participants_true = true,
n_participants_false = false) |>
select(lp, true, false)

attrition |>
summarise(n_trials_true = sum(is_valid_trial),
n_trials_false = sum(!is_valid_trial),
.by = c(lp, trial_type, location)) |>
arrange(desc(lp)) |>
pivot_wider(id_cols = c(location, lp),
names_from = c(trial_type),
values_from = matches("trials"),
names_repair = janitor::make_clean_names) |>
mutate(study = if_else(location=="Oxford", 
"Study 1: English (Oxford)",
"Study 2: Catalan/Spanish (Barcelona)")) |> 
left_join(n_lp, by = join_by(lp)) |>
select(study, lp, true, false, matches("n_trials")) |>
gt(rowname_col = "lp",
groupname_col = "study") |>
tab_spanner("Participants", matches("participant")) |>
tab_spanner("Trials", matches("trial")) |>
fmt_integer(everything()) |>
cols_merge(c(true, false), pattern = "{1} ({2})") |>
cols_merge(c(n_trials_true_cognate, n_trials_false_cognate),
pattern = "{1} ({2})") |>
cols_merge(c(n_trials_true_non_cognate, n_trials_false_non_cognate),
pattern = "{1} ({2})") |>
cols_merge(c(n_trials_true_unrelated, n_trials_false_unrelated),
pattern = "{1} ({2})") |>
cols_label(true = "Participants",
n_trials_true_cognate = "Cognate",
n_trials_true_non_cognate = "Non-cognate",
n_trials_true_unrelated = "Unrelated") |>
tab_style(cell_text(style = "italic"),
cells_column_labels()) |>
tab_style(cell_text(align = "left"),
cells_stub()) |>
grand_summary_rows(fns = list(label = md("*N*")) ~ sum(.),
fmt = ~ fmt_integer(.))
```

@@

```{r coefs-oxf}
var_labels <- c(
"b_Intercept" = "Intercept",
"b_age_std" = "Age (+1 SD)",
"b_condition1" = "Condition (U vs. RN)",
"b_condition2" = "Condition (RN vs. RC)"
)

model_summary <- get_posterior_summary(model_fits_oxf$fit_oxf_1,
data_bcn,
vars_dict = var_labels,
rope_interval = rope_interval
)

@@

```{r tbl-coefs-oxf}
#| label: tbl-coefs-oxf
#| fig-cap: "Margin posterior mean predictions. A) Posterior mean predictions of the time course of target fixation in the test phase. B) Posterior mean prediction of the time course of the differences in target looking time between conditions. Lines and intervals indicate the mean and 95% credible interval of the posterior predictions."
#| echo: false
#| eval: false
#| message: false
#| warning: false
#| out-width: 80%

# summarise posterior draws
intercept_label <- paste0("Intercept (at ", round(mean(data_oxf$age), 2), " months)")

tbl_data <- model_summary |>
	select(.variable_name, .value, .lower, .upper, .rope, .type) |>
	mutate(
		.variable_name = gsub("×", "$\\\\times$", .variable_name),
		.hdi = sprintf("[%.3f, %.3f]", .lower, .upper),
		.rope = case_when(is.na(.rope) ~ "-",
						  .rope > 0.9999 ~ "1",
						  .rope < 0.00001 ~ "0",
						  .default = sub("^(-?)0.", "\\1.", sprintf("%.3f", .rope))
		)
	) |>
	select(.variable_name, .value, .lower, .upper, .hdi, .rope)

col_names <- c("", "$\\beta$", "95\\% HDI", "$p(\\text{ROPE})$")

tbl_data |>
	select(-c(.lower, .upper)) |>
	kable(
		format = "latex",
		escape = FALSE,
		col.names = col_names,
		digits = 3,
		booktabs = TRUE,
		align = "lclr",
		linesep = ""
	) |>
	kable_styling() |>
	add_footnote("$\\beta$: median of the posterior distribution in the probability scale. 95\\% HDI: 95\\% highest density interval of the distribution. $p(\\text{ROPE})$: overlap between the 95\\% HDI and the ROPE, indicating the posterior probability that the true value of the coefficient is equivalent to zero.", notation = "none", escape = FALSE)
```

@@

plot_epreds <- epreds |>
	ggplot(aes(timebin_std, .epred,
			   colour = condition,
			   fill = condition,
			   linetype = condition,
			   shape = condition)) +
	geom_hline(yintercept = 0,
			   colour = "grey") +
	# geom_line(aes(group = interaction(condition, .draw)),
	# 		  alpha = 1/20,
	# 		  linewidth = 3/4,
	# 		  linetype = "solid") +
	stat_summary(fun.data = mean_qi,
				 geom = "ribbon",
				 alpha = 1 / 4,
				 linewidth = NA) +
	stat_summary(fun = "mean",
				 geom = "line",
				 # colour = "black",
				 linewidth = 3 / 4) +
	stat_summary(data = data_oxf,
				 aes(y = .elog),
				 fun = "mean",
				 geom = "point",
				 size = 2) +
	labs(x = "Time (ms)",
		 y = "Logit (Target looking)",
		 colour = "Condition",
		 fill = "Condition",
		 linetype = "Condition",
		 shape = "Condition")

@@

```{r tbl-loos-oxf}
#| label: tbl-loos-oxf
#| tbl-cap: "Leave-one-out cross validation outcomes, comparing the predictive performance of the models in Study 2."
loo_compare(model_loos_oxf[2:4]) |>
	as.data.frame() |>
	rownames_to_column("model") |>
	mutate(model = paste0("Model ",
						  as.numeric(gsub("fit_oxf_", "", model)) - 1),
		   across(matches("diff"),
		   	   \(x) {
		   	   	x[1] <- NA
		   	   	return(x)
		   	   }),
		   model = case_when(
		   	model == "Model 1" ~ "Model 1 (Age)",
		   	model == "Model 2" ~ "Model 2 (L1 vocabulary)",
		   	model == "Model 3" ~ "Model 3 (Total vocabulary)",
		   	.default = model
		   )) |>
	select(model, matches("elpd|diff")) |>
	relocate(matches("diff"), .after = matches("loo")) |>
	gt(rowname_col = "model") |>
	fmt_number(is.numeric) |>
	cols_label(elpd_diff = "ELPD (diff.)",
			   se_diff = "SE ELPD (diff)",
			   elpd_loo = "ELPD",
			   se_elpd_loo = "SE ELPD") |>
	fmt_missing(missing_text = "---") |>
	tab_footnote(footnote = md("*ELPD*: sum of expected log pointwise predictive density for a new dataset. *SE \\(ELPD\\)*: standard error of the *ELPD*, which indictes the uncertainty about the predictive performance for unknown future data. *ELPD (diff)*: pairwise difference in *ELPD* for two models. The difference is computed relative to the model with lowest *ELPD* (best fitting model). *SE ELPD (diff.)*: standard error of component-wise differences of *ELPD* between two models.")) |>
	tab_style(cell_text(style = "italic"), cells_column_labels())
```

@@

```{r dataset}
attrition_bcn <- filter(attrition, location == "Barcelona")
n_trials <- nrow(attrition_bcn)
n_sessions <- n_distinct(attrition_bcn$session_id)
n_participants <- n_distinct(attrition_bcn$child_id)
n_trials_valid <- nrow(filter(
	attrition_bcn,
	is_valid_participant,
	is_valid_trial
))
n_sessions_valid <- sum(distinct(attrition_bcn,
								 session_id,
								 .keep_all = TRUE
)$is_valid_participant)
n_participants_valid <- n_distinct(
	attrition_bcn[attrition_bcn$is_valid_participant, ]$child_id
)

n_exc_prime <- sum(!attrition_bcn$is_valid_gaze_prime)
n_exc_test <- sum(!attrition_bcn$is_valid_gaze_test)
n_exc_test_each <- sum(!attrition_bcn$is_valid_gaze_test_each)
n_exc_participants <- sum(!distinct(attrition_bcn,
									session_id,
									.keep_all = TRUE
)$is_valid_participant)
n_exc_participants_vocab <- sum(!distinct(attrition_bcn,
										  session_id,
										  .keep_all = TRUE
)$is_valid_vocab_size)

@@

After excluding trials that matched any of the aforementioned criteria from the dataset, we excluded participants who did not provide at least two valid trials in the *Cognate* condition, the *Non-cognate* condition, or the *Unrelated* condition (*n* = `r n_exc_participants`), and participants with a L1 vocabulary size lower than 10% (*n* = `r n_exc_participants_vocab`). The final dataset included `r format(n_trials_valid, big.mark = ",")` trials from `r n_sessions_valid` testing sessions, generated by `r format(n_participants_valid, big.mark = ",")` distinct participants. Of those participants, `r n_longitudinal[1, 2]` provided data from one experimental session, `r n_longitudinal[2, 2]` provided data from two experimental sessions, and `r n_longitudinal[3, 2]` provided data from three experimental sessions. @tbl-attrition shows a detailed description of the trial attrition.

@@

```{r tbl-coefs}
#| label: tbl-coefs
#| tbl-cap: "Posterior distribution of regression coefficients. Median: median of the posterior distribution in logit scale. 95\\% HDI: 95\\% highest density interval of the distribution. *p*(ROPE): overlap between the 95\\% HDI and the ROPE, indicating the posterior probability that the true value of the coefficient is equivalent to zero."
#| echo: false
#| message: false
#| eval: false
#| warning: false
#| out-width: 80%
# summarise posterior draws
intercept_label <- paste0("Intercept (at ", round(mean(data_bcn$age), 2), " months)")

tbl_data <- model_summary |>
	select(.variable_name, .value, .lower, .upper, .rope, .type) |>
	mutate(
		.variable_name = gsub("×", "$\\\\times$", .variable_name),
		.hdi = sprintf("[%.3f, %.3f]", .lower, .upper),
		.rope = case_when(is.na(.rope) ~ "-",
						  .rope > 0.9999 ~ "1",
						  .rope < 0.00001 ~ "0",
						  .default = sub("^(-?)0.", "\\1.", sprintf("%.3f", .rope))
		)
	) |>
	select(.variable_name, .value, .lower, .upper, .hdi, .rope)

col_names <- c("", "$\\beta$", "95\\% HDI", "$p(\\text{ROPE})$")

tbl_data |>
	select(-c(.lower, .upper)) |>
	kable(
		format = "latex",
		escape = FALSE,
		col.names = col_names,
		digits = 3,
		booktabs = TRUE,
		align = "lclr",
		linesep = ""
	) |>
	kable_styling() |>
	add_footnote("$\\beta$: median of the posterior distribution in the logit scale. 95\\% HDI: 95\\% highest density interval of the distribution. $p(\\text{ROPE})$: overlap between the 95\\% HDI and the ROPE, indicating the posterior probability that the true value of the coefficient is equivalent to zero.", notation = "none", escape = FALSE)
```

@@

```{r tbl-loos}
#| label: tbl-loos
#| tbl-cap: "Leave-one-out cross validation outcomes, comparing the predictive performance of models 0 to 3, and 3 to 6 in Study 1."
get_loos <- function(x, y) {
	out <- lst(m_0 = x, m_1 = y) |>
		loo_compare() |>
		as.data.frame() |>
		rownames_to_column("model") |>
		arrange(model) |>
		select(model, elpd_diff, se_diff) |>
		mutate(
			diff_label = sprintf("%.3f", elpd_diff),
			se_label = sprintf("%.3f", se_diff)
		)
	return(out)
}

>>>>

```{r setup}
#| label: setup
#| echo: false
#| message: false
#| warning: false
# load objects
targets::tar_config_set(store = here::here("_targets"),
						script = here::here("_targets.R"))

@@

targets::tar_load(c(participants,
					stimuli,
					vocabulary,
					attrition_trials,
					attrition_participants,
					bvq_data,
					gaze,
					data_bcn,
					data_oxf))

targets::tar_load(c(model_fits_bcn,
					model_loos_bcn,
					model_fits_oxf,
					model_loos_oxf))

# prepare data
attrition <- lst(participants,
				 attrition_participants,
				 attrition_trials) |>
	reduce(inner_join, by = join_by(session_id)) |>
	unnest_wider(matches("all"), names_sep = "_") |>
	rename_with(\(x) gsub("is_valid_gaze_all_|is_valid_vocab_all", "", x))

library(ggplot2)
library(gt)
library(patchwork)
library(english)

# set ggplot theme and colour palette
my_theme <- theme_minimal() +
	theme(panel.grid = element_blank(),
		  axis.line = element_line(colour = "black"),
		  text = element_text(size = 12, colour = "black"),
		  axis.text = element_text(colour = "black"))

@@

clrs <- c("#c8102e", "#1e90ff", "grey50")

options(ggplot2.ordinal.fill = clrs,
		ggplot2.ordinal.colour = clrs,
		ggplot2.discrete.fill = clrs,
		ggplot2.discrete.colour = clrs,
		ggplot2.continuous.fill = ggplot2::scale_color_gradient,
		ggplot2.continuous.colour = ggplot2::scale_color_gradient)

@@

	cols_merge(c(age_min, age_max), pattern = "{1}-{2}") |>
	cols_merge(c(age_mean, age_sd),
			   pattern = "{1} ({2})") |>
	cols_merge(c(doe_catalan_mean, doe_catalan_sd),
			   pattern = "{1} ({2})") |>
	cols_merge(c(doe_english_mean, doe_english_sd),
			   pattern = "{1} ({2})") |>
	cols_merge(c(doe_spanish_mean, doe_spanish_sd),
			   pattern = "{1} ({2})") |>
	tab_spanner("Sample size", starts_with("n")) |>
	tab_spanner("Age (months)", matches("age")) |>
	tab_spanner("Catalan", matches("catalan")) |>
	tab_spanner("English", matches("english")) |>
	tab_spanner("Spanish", matches("spanish")) |>
	tab_spanner("Degree of Exposure (%)", matches("doe")) |>
	cols_label(n = md("*Sessions*"),
			   n_participants = md("*Participants*"),
			   age_mean = md("*M (SD)*"),
			   age_min = md("*Range*"),
			   doe_catalan_mean = md("*M (SD)*"),
			   doe_english_mean = md("*M (SD)*"),
			   doe_spanish_mean = md("*M (SD)*"),) |>
	tab_style(cell_text(style = "italic"),
			  cells_column_labels()) |>
	tab_style(cell_text(style = "italic"),
			  cells_column_spanners())
```

@@

```{r fig-vocabulary}
#| label: fig-vocabulary-oxf
#| fig-cap: "Participant receptive vocabulary sizes across ages and language profiles. Vocabulary size scores from the recurrent participants are shown linked. For visualisation purposes, testing sessions were aggregated at three age points: 21, 25, and 30 months. Points and intervals indicate mean and standard deviations, respectively."
#| fig-width: 7
#| fig-height: 5.5
vocab_plot_data <- vocabulary |>
	inner_join(filter(attrition_participants, is_valid_participant),
			   by = join_by(session_id)) |>
	inner_join(select(participants, location, lp, 
					  child_id, session_id, age_group),
			   by = join_by(child_id, session_id)) |>
	pivot_longer(c(l1_prop, total_prop),
				 names_to = "type",
				 values_to = "prop") |> 
	mutate(location = factor(location, levels = c("Oxford", "Barcelona")),
		   lp = gsub(" \\(English\\)", "", lp),
		   side = case_when(location=="Oxford" ~ "both",
		   				 lp=="Monolingual" ~ "left",
		   				 lp=="Bilingual" ~ "right"),
		   type = if_else(type=="l1_prop", 
		   			   "L1 vocabulary", "Total vocabulary"),
		   study = if_else(location=="Oxford", "Study 1", "Study 2"),
		   age_group = gsub(" months", "", age_group))

vocab_plot_means <- vocab_plot_data |> 
	summarise(prop_sd = sd(prop, na.rm = TRUE),
			  prop = mean(prop, na.rm = TRUE),
			  n = n(),
			  .by = c(age_group, type, side, lp, type, study)) |> 
	mutate(.lower = prop - (prop_sd/sqrt(n)),
		   .upper = prop + (prop_sd/sqrt(n)),
		   age_group = gsub(" months", "", age_group))

vocab_plot_data |> 
	ggplot(aes(age_group, prop, 
			   colour = lp,
			   fill = lp,
			   side = side)) +
	facet_grid(type~study) +
	geom_line(aes(group = child_id),
			  # colour = "grey",
			  alpha = 1 / 4,
			  linewidth = 1 / 2) +
	geom_dots(dotsize = 1.5,
			  scale = 0.25,
			  binwidth = 0.025,
			  alpha = 2 / 4) +
	geom_errorbar(data = vocab_plot_means,
				  aes(ymin = .lower,
				  	ymax = .upper),
				  colour = "black",
				  show.legend = FALSE,
				  width = 0.15,
				  linewidth = 1/2,
				  position = position_dodge(width = 1)) +
	geom_point(data = vocab_plot_means, 
			   colour = "black",
			   show.legend = FALSE,
			   size = 1.25,
			   position = position_dodge(width = 1)) +
	labs(x = "Age (months)",
		 y = "Vocabulary size",
		 colour = "Group",
		 fill = "Group") +
	scale_y_continuous(labels = scales::percent,
					   limits = c(0, 1)) +
	theme(panel.grid.major.y = element_line(linetype = "dotted",
											colour = "grey"),
		  legend.position = "top",
		  legend.title = element_blank())
```

@@

We created fours lists of trials, across which the same target-distractor pair appeared with a different prime, counterbalancing the condition to which it belonged. For instance, in list one the *ball*-*trousers* pair was preceded by *bike* (*Related/Cognate*), by *butterfly* (*Related/Non-cognate*) in list two, and by *star* and *nose* (*Unrelated*) in lists three and four (see Appendix A for a detailed description of the stimuli). Lexical frequencies were extracted from the English corpora from the CHILDES database [@macwhinney2000childes; @sanchez2019childes] as counts per million words, and transformed into Zipf scores for easier cross-language comparison [@zipf1945meaning; @van2014subtlex].

@@

After trials that matched any of the aforementioned exclusion criteria from the dataset, we excluded participants who did not provide at least two valid trials in the *Cognate* condition, the *Non-cognate* condition, or the *Unrelated* condition (*n* = `r n_exc_participants`), and participants with a L1 vocabulary size lower than 10% (*n* = `r n_exc_participants_vocab`). The final dataset included `r format(n_trials_valid, big.mark = ",")` trials from `r n_sessions_valid` testing sessions, generated by `r format(n_participants_valid, big.mark = ",")` distinct participants. Of those participants, `r n_longitudinal[1, 2]` provided data from one experimental session, and `r n_longitudinal[2, 2]` provided data from two experimental sessions. From the trials included in the final dataset dataset, `r n_trials_condition$unrelated_2$n` were *Unrelated* trials (`r n_trials_condition$unrelated$n` previously excluded), were `r n_trials_condition$non_cognate_2$n` *Related/Non-cognate* trials(`r n_trials_condition$non_cognate$n` previously excluded), and `r n_trials_condition$cognate_2$n` were *Related/Cognate* trials (`r n_trials_condition$cognate$n` previously excluded).

@@

```{r coefs-oxf}
var_labels <- c("b_Intercept" = "Intercept",
				"b_age_std" = "Age (+1 SD)",
				"b_condition1" = "Condition (U vs. RN)",
				"b_condition2" = "Condition (RN vs. RC)")

model_summary <- get_posterior_summary(model_fits_oxf$fit_oxf_1,
									   data_bcn,
									   vars_dict = var_labels,
									   rope_interval = rope_interval)

coefs <- model_summary |>
	mutate(.value = sprintf("%.3f", .value),
		   .ci = sprintf("[%.3f, %.3f]", .lower, .upper),
		   .rope = scales::percent(.rope, accuracy = 0.01)) |>
	split(janitor::make_clean_names(model_summary$.variable))
```

@@

@@

plot_epreds <- epreds |>
	ggplot(aes(timebin_std, .epred,
			   colour = condition,
			   fill = condition,
			   linetype = condition,
			   shape = condition)) +
	geom_hline(yintercept = 0,
			   colour = "grey") +
	stat_summary(fun.data = mean_qi,
				 geom = "ribbon",
				 alpha = 1 / 4,
				 linewidth = NA) +
	stat_summary(fun = "mean",
				 geom = "line",
				 # colour = "black",
				 linewidth = 3 / 4) +
	stat_summary(data = data_oxf,
				 aes(y = .elog),
				 fun = "mean",
				 geom = "point",
				 size = 2) +
	labs(x = "Time (ms)",
		 y = "Logit (Target looking)",
		 colour = "Condition",
		 fill = "Condition",
		 linetype = "Condition",
		 shape = "Condition")

@@

```{r tbl-loos-oxf}
#| label: tbl-loos-oxf
#| tbl-cap: "Leave-one-out cross validation outcomes, comparing the predictive performance of the models in Study 1."
loo_compare(model_loos_oxf[2:4]) |>
	as.data.frame() |>
	rownames_to_column("model") |>
	mutate(model = paste0("Model ",
						  as.numeric(gsub("fit_oxf_", "", model)) - 1),
		   across(matches("diff"),
		   	   \(x) {
		   	   	x[1] <- NA
		   	   	return(x)
		   	   }),
		   model = case_when(
		   	model == "Model 1" ~ "Model 1 (Age)",
		   	model == "Model 2" ~ "Model 2 (L1 vocabulary)",
		   	model == "Model 3" ~ "Model 3 (Total vocabulary)",
		   	.default = model
		   )) |>
	select(model, matches("elpd|diff")) |>
	relocate(matches("diff"), .after = matches("loo")) |>
	gt(rowname_col = "model") |>
	fmt_number(is.numeric) |>
	cols_label(elpd_diff = "ELPD (diff.)",
			   se_diff = "SE ELPD (diff)",
			   elpd_loo = "ELPD",
			   se_elpd_loo = "SE ELPD") |>
	fmt_missing(missing_text = "---") |>
	tab_footnote(footnote = md("*ELPD*: sum of expected log pointwise predictive density for a new dataset. *SE \\(ELPD\\)*: standard error of the *ELPD*, which indictes the uncertainty about the predictive performance for unknown future data. *ELPD (diff)*: pairwise difference in *ELPD* for two models. The difference is computed relative to the model with lowest *ELPD* (best fitting model). *SE ELPD (diff.)*: standard error of component-wise differences of *ELPD* between two models.")) |>
	tab_style(cell_text(style = "italic"), cells_column_labels())
```

@@

```{r dataset}
attrition_bcn <- filter(attrition, location == "Barcelona")
n_trials <- nrow(attrition_bcn)
n_sessions <- n_distinct(attrition_bcn$session_id)
n_participants <- n_distinct(attrition_bcn$child_id)
n_trials_valid <- nrow(filter(
	attrition_bcn,
	is_valid_participant,
	is_valid_trial
))
n_sessions_valid <- sum(distinct(attrition_bcn,
								 session_id,
								 .keep_all = TRUE
)$is_valid_participant)
n_participants_valid <- n_distinct(
	attrition_bcn[attrition_bcn$is_valid_participant, ]$child_id)

n_exc_prime <- sum(!attrition_bcn$is_valid_gaze_prime)
n_exc_test <- sum(!attrition_bcn$is_valid_gaze_test)
n_exc_test_each <- sum(!attrition_bcn$is_valid_gaze_test_each)
n_exc_participants <- sum(!distinct(attrition_bcn,
									session_id,
									.keep_all = TRUE)$is_valid_participant)
n_exc_participants_vocab <- sum(!distinct(attrition_bcn,
										  session_id,
										  .keep_all = TRUE
)$is_valid_vocab_size)

@@

After excluding trials that matched any of the aforementioned criteria from the dataset, we excluded participants who did not provide at least two valid trials in the *Cognate* condition, the *Non-cognate* condition, or the *Unrelated* condition (*n* = `r n_exc_participants`), and participants with a L1 vocabulary size lower than 10% (*n* = `r n_exc_participants_vocab`). The final dataset included `r format(n_trials_valid, big.mark = ",")` trials from `r n_sessions_valid` testing sessions, generated by `r format(n_participants_valid, big.mark = ",")` distinct participants. Of those participants, `r n_longitudinal[1, 2]` provided data from one experimental session, `r n_longitudinal[2, 2]` provided data from two experimental sessions, and `r n_longitudinal[3, 2]` provided data from three experimental sessions.

@@

@@

```{r tbl-loos}
#| label: tbl-loos
#| tbl-cap: "Leave-one-out cross validation outcomes, comparing the predictive performance of the models in Study 1."
get_loos <- function(x, y) {
	out <- lst(m_0 = x, m_1 = y) |>
		loo_compare() |>
		as.data.frame() |>
		rownames_to_column("model") |>
		arrange(model) |>
		select(model, elpd_diff, se_diff) |>
		mutate(
			diff_label = sprintf("%.3f", elpd_diff),
			se_label = sprintf("%.3f", se_diff)
		)
	return(out)
}
